---
title: "Lab 07 - Conveying the right message through visualisation"
author: "Saima Arina"
date: "02/24/2026"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```


Looking at the visualization, I can see how misleading it is. First of all, the y-axes for both sides are on a different scale (the mask side goes from 15-25 while the no mask side goes from 14-4, and this is extremely confusing). Because of that, it is easy to think before actually diving deeper that mask-mandate counties may have shown a decrease in cases, but no mask-mandate counties did not seem to change (which can insinuate that not wearing a mask does not have consequences). If you actually look closer, you can see that there actually is a decrease in cases for the no mask-mandate counties. Also, this could all be avoided if the visualization was made with standards kept in mind, like the scale to be the same for both variables and a proper color legend to better discriminate between the two. 

### Exercise 1

```{r - loading file}
library(readr)
df <- read_csv("kansas_grouped_rolling_avg.csv")
View(df)
```

### Exercise 2

```{r - enhanced and corrected visual}
df %>% 
  mutate(mask_group = if_else(mask_mandate == "Mask",
                              "Mask mandate counties",
                              "No mask mandate counties")) %>% 
  ggplot(aes(x = date,
             y = rolling_avg,
             color = mask_group,
             group = mask_group)) +
  geom_line(size = 1.1) +
  scale_color_manual(values = c("Mask mandate counties" = "red",
                                "No mask mandate counties" = "blue"),
                     name = "County type") +
  labs(
    x = "Date",
    y = "7-day rolling average of new cases\n(per 100,000 residents)",
    title = "COVID-19 case trends in Kansas counties\nwith vs. without mask mandates",
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank()
  )

```

Looking at the data now that it is visually represented better, I can see how even my initial assumptions were wrong after I thought I looked closer, but it wasn't close enough! It just makes no sense why the plot would be made the way it did when it could be made so much simpler, that is so frustrating!


### Exercise 3

In my visualization, it is way more clear that there were not really any differences in the number of cases that were notable in the no mask-mandate counties, while the decrease in cases for the mask-mandate counties are much clearer, showcasing their effect more strongly than in the original visual representation. 


